---
title: "Lab 1"
output:
  pdf_document: default
  html_notebook: default
---
First I am creating a data fram reading in other data sets and adding there data to the data fram below 
```{r}
library(dplyr)
library(countrycode)
library(tinytex)

pop <- read.csv("pop.csv")
country = pop$country
p2010 = pop$X2010
# colnames(pop)
life_exp <- read.csv("lex.csv")

Life2010 <- life_exp$X2010
life_cont <- life_exp[, c("country", "X2010")]
#print(life_cont)
# do country names to then merge it with the df

df <- data.frame(
  country = country,                            
  population = rep(0, length(country)), 
  life_expectancy = rep(0, length(country)), 
  daily_income = rep(0, length(country)),  
  babies_per_woman = rep(0, length(country)),  
  child_mortality = rep(0, length(country)),
  co2_emissions_per_person = rep(0, length(country)),
  gdp_per_capita = rep(0, length(country)), 
  health_spending_per_person = rep(0, length(country)),     
  population_density = rep(0, length(country)),   
  basic_water_source_percentage = rep(0, length(country)),   
  murders = rep(0, length(country)), 
  continent = rep(0, length(country)),
  stringsAsFactors = FALSE                      
)
# convert population M and B
text.to.num <- function(x){
xx = as.character(x)
res=rep(0, length(x))
for(i in 1:length(x)){
if(is.na(xx[i])){
res[i]=NA
next
}
if (grepl("M", xx[i], ignore.case = TRUE)) {
res[i]=as.numeric(gsub("M", "", xx[i], ignore.case = TRUE)) * 1e6
} else if (grepl("k", xx[i], ignore.case = TRUE)) {
res[i]=as.numeric(gsub("k", "", xx[i], ignore.case = TRUE)) * 1e3
} else {
res[i]=as.numeric(xx[i])
}
}
res
}

#country code 
df$continent <- countrycode(sourcevar = df[, "country"],
origin = "country.name",
destination = "continent")
  

# ading columns 
df$population <- p2010
df$population <- text.to.num(df$population)
df$population <- text.to.num(df$population)

#Life
df <- merge(df, life_cont, by = "country", all.x = TRUE)
df$life_expectancy <- df$X2010
df$X2010 <- NULL
#child 
child <- read.csv("child.csv")
cd2010 <- child[, c("country", "X2010")]
df <- merge(df, cd2010, by = "country", all.x = TRUE)
df$child_mortality <- df$X2010
df$X2010 <- NULL

#income 
income <- read.csv("income.csv")
income_2010 <- income[, c("country", "X2010")]
df <- merge(df, income_2010, by = "country", all.x = TRUE)
df$daily_income <- df$X2010
df$X2010 <- NULL

#babies per women 
babies <- read.csv("CPW.csv")
b2010 <- babies[, c("country", "X2010")]
df <- merge(df, b2010, by = "country", all.x = TRUE)
df$babies_per_woman <- df$X2010
df$X2010 <- NULL

# co2 emissions 
co2 <- read.csv("co2.csv")
co2010 <- co2[, c("country", "X2010")]
df<- merge(df, co2010, by = "country", all.x = TRUE)
df$co2_emissions_per_person <- df$X2010
df$X2010 <- NULL

# gdp 
gdp <- read.csv("gdp.csv")
gd2010 <- gdp[, c("country","X2010")]
df <-  merge(df, gd2010, by = "country", all.x = TRUE)
df$gdp_per_capita <- df$X2010
df$X2010 <- NULL
df$gdp_per_capita <- text.to.num(df$gdp_per_capita)

# spending per person
hm <- read.csv("health_money.csv")
spending <- hm[, c("country", "X2010")]
df <- merge(df , spending, by = "country", all.x = TRUE)
df$health_spending_per_person <- df$X2010
df$X2010 <- NULL

#population density 
popd <- read.csv("popd.csv")
popd2010 <- popd[, c("country", "X2010")]
df <- merge(df, popd2010, by = "country", all.x= TRUE)
df$population_density <- df$X2010
df$X2010 <- NULL

#water supply 
water <- read.csv("water.csv")
water2010<- water[, c("country", "X2010")]
df <-  merge(df, water2010, by = "country", all.x = TRUE)
df$basic_water_source_percentage <- df$X2010
df$X2010 <- NULL

# murder 
mur <- read.csv("murder.csv")
dead2010 <- mur[, c("country", "X2010")]
df<- merge(df, dead2010, by = "country", all.x = TRUE)
df$murders <- df$X2010
df$X2010 <- NULL

glimpse(df)

```

```{r}
# Scatter plot of life expectancy vs childmortality 
library(ggplot2)
ggplot(data = df, aes(x = child_mortality, y = life_expectancy)) +
  geom_point() + 
  labs(title = "Life Expectancy vs. Child Mortality",
       x = "Child Mortality ",
       y = "Life Expectancy ") +
  theme_minimal()


```
For the above graph the correlation between the graphs shows the lower the child mortality the higher the life expectancy is. Same with the opposite, the higher the child mortality the lower the life expectancy. So we can conclude there is a relatinship between child mortality and life expectancy where the lower the child mortality the higher the life expectancy. This scatter plot therefore has a negative relationship and looks to be approximately linear. 
```{r}
#2 life expectancy and child mortality in color
ggplot(data = df, aes(x= child_mortality, y= life_expectancy, color= continent, size=population)) + geom_point()+ labs( x = "child_mortality" , y = "life_expectancy") + theme_minimal()


```
Similar to the relationship discussed above the graph shows that Africa has a higher child mortality rate and therefore a lower life expectancy. Where as continents like the Americas and Europe have a higher life expectancy and a lower child mortality rate. This graph can help break the relaitonship by countries so we can see other than the one outlier in the Americas the trend by Countries. 


```{r}
#3 this is a Histogram of life expectancy

ggplot(data = df, aes(x = life_expectancy))+ geom_histogram(fill="blue", color="black")+ labs(x = "life expectancy", y= "count")

```
explanation : This data is skewed to the left meaning most of the life expectancy are higher. When I say higher I mean 60 and above with a lone outlier being around 35.

```{r}
# 4 box plot of life expectancy 
ggplot(data = df, aes(y = life_expectancy)) + geom_boxplot( fill = "lightblue", color = "black" ) + labs(y = "Life expectancy") + theme_minimal()



```
There is a standalone dot thats not included in the box or on the line, the dot is around 30 and is far to the left because that country has a much shorter life expectancy than the others. But the box plot also shows the average life expectancy is around 70 years of age. 
```{r}
# #5box plot by continent 
ggplot(data= df, aes(x = life_expectancy, y = continent))+geom_boxplot()+labs(title = "Life Expectancy by Continent", x= "Continent", y = "Life Expectancy") + theme_minimal()



```
the plots remain similar, but African has a significantly lower average life expectancy than the other continents. Europe has the highest life expectancy well other continents like Asia and the Americas have similar means and ranges. We can also see there are 2 negative outliers for the Americas and 2 positive outliers for Oceania.  
```{r}
library(dplyr)
c_counts <- df %>%
count(continent) 
print(c_counts)
  
  # # 6 create plot to show numbver of countries
ggplot(data= c_counts, aes(x= continent, y=n)) + geom_bar(stat = "identity")+ labs(title = "Country Count Per Continent")+ theme_minimal()
  

```
This plot below shows that Africa and Asia have the highest country counts well continents like Oceania have singificantly lower counts which could affect the different stats, like murders, income, money per child, and more. 
Africa has the most countries and Asia has the second most countries. 
  
```{r}
#define an indicator of weather the avergae number of women per children is > 2
# 7 and 8
baby_c <- df %>%
  mutate(child_i = ifelse(babies_per_woman <= 2, 1 ,0))
# print(baby_c)
ggplot(data = baby_c, aes(x = continent, fill = factor(child_i))) + geom_bar(position="fill") + labs(title = "Women with 2 or more children or without", x = "Countries")+ theme_classic()

```
The percetnages of children <= 2 are not very similar, Europe and ociania have smaller smaller averages/ counts well other countries like Africa Asia and the Americas have more counts with less than 2 babies per women. A trend we can see is that most countries aren't very even(balanced). They either have more than 2 children majority of the time or less than 2 children majority of the time. there is  very little overlap. 
